---
title: "IOin_examples"
author: "Will Burke"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RHESSysIOinR Input Functions

This vignette shows example uses of the IOin_X functions, which together construct input data objects used to run one or many RHESSys simulations.

New/renamed input functions:

 * IOin_rhessys_input
 * IOin_tec_std
 * IOin_std_pars
 * IOin_hdr
 * IOin_clim
 * IOin_def_pars_simple

### IOin_rhessys_input

```{r eval=FALSE}

input_rhessys = IOin_rhessys_input(
  version = file.path("bin", comp_ID, "rhessys7.2"),
  tec_file = "tecfiles/BC.tec",
  world_file = "worldfiles/no_multi/BC_conifer_S_2p.world",
  world_hdr_prefix = "BC",
  flowtable = "flowtables/BC_2p.flow",
  start = "1943 10 01 01",
  end = "1948 10 01 01",
  output_folder = "output/",
  output_prefix = "BC_2p",
  commandline_options = c("-vmort_off -g -b -p -c -msr -v -6")
)

```

### IOin_tec_std

```{r eval=FALSE}

tec_std = IOin_tec_std(start = "1943 10 01 01",
                       end = "1948 10 01 01",
                       output_state = TRUE
)

```

### IOin_std_pars

```{r eval=FALSE}

std_pars = IOin_std_pars(
  m = 2,
  k = 2,
  m_v = 2,
  k_v = 2,
  pa = 1.15,
  po = 0.766,
  gw1 = 0.24,
  gw2 = 0.2
)

```

### IOin_hdr

```{r eval=FALSE}

input_hdr_list = IOin_hdr(basin = "defs/basin_p301.def",
                          hillslope = "defs/hill_p301.def",
                          zone = "defs/zone_p301.def",
                          soil = "defs/soil_forestshrub.def",
                          landuse = "defs/lu_p301.def",
                          stratum = c("defs/veg_p301_shrub.def", "defs/veg_p301_conifer.def"),
                          basestations = "clim/upGGmod.base")

```

### Run RHESSys

Together, these inputs can run RHESSys for a single simulation with very simple options.

```{r eval=FALSE}

# this is a nearly barebones run
run_rhessys_core(
  input_rhessys = input_rhessys,
  hdr_files = input_hdr_list,
  std_pars = input_std_pars,
  tec_data = input_tec_data
)

```

### IOin_clim

```{r eval=FALSE}

# this is a very simple case, just creating a basestation
input_clim = IOin_clim(
  base_station_id = 101,
  x_coordinate = 100.0,
  y_coordinate = 100.0,
  z_coordinate = 1984,
  effective_lai = 3.5,
  screen_height = 160,
  daily_prefix = "clim/upGGmod"
)

```

### IOin_def_pars

```{r eval=FALSE}

input_def_pars = IOin_def_pars_simple(
  list("defs/veg_p301_conifer.def", "epc.allocation_flag","dickenson"),
  list("defs/veg_p301_conifer.def", "epc.alloc_frootc_leafc", 1),
  list("defs/veg_p301_conifer.def", "epc.alloc_stemc_leafc", 0.6),
  list("defs/veg_p301_conifer.def", "epc.netpabs_shade", 0.2),
  list("defs/soil_forestshrub.def", "soil_depth", 1),
  list("defs/soil_forestshrub.def", "pore_size_index", 0.15),
  list("defs/soil_forestshrub.def", "sat_to_gw_coeff", 0.12),
  list("defs/soil_forestshrub.def", "psi_air_entry", 0.15)
)

```

### TODO: Output vars

```{r eval=FALSE}

  output_method = "r"

  # this will be turned into a function
  # patch/canopy level vars
  output_variables <- data.frame(out_file = character(), variable = character(), stringsAsFactors = FALSE)
  output_variables[1,] = c("pd", "Qout")
  output_variables[2,] <- c("pd", "litter.S") 
  output_variables[3,] <- c("pd", "psn")
  output_variables[4,] <- c("pdg", "lai")
  output_variables[5,] <- c("pdg", "plantc")
  output_variables[6,] <- c("cd", "height")
  output_variables[7,] <- c("cdg", "leafc") 
  output_variables[8,] <- c("pdg", "litrc_sum") 
  output_variables[9,] <- c("pd", "rz_storage") 
  output_variables[10,] <- c("pd", "fire_et")
  output_variables[11,] <- c("pd", "pet")

```


### Run RHESSys (v2)

This runs RHESSys for a single simulation with a few more options.

```{r eval=FALSE}

run_rhessys_core(
  input_rhessys = input_rhessys,
  hdr_files = input_hdr_list,
  def_pars = input_def_pars,
  std_pars = input_std_pars,
  tec_data = input_tec_data,
  clim_base = input_clim,
  output_method = output_method,
  output_variables = output_vars,
  
)

```
