---
title: "IOin_examples"
author: "Will Burke"
date: "9/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RHESSysIOinR Input Functions

This vignette shows example uses of the IOin\_X functions, which together construct input data objects used to run one or many RHESSys simulations.

#### New/renamed input functions:

-   IOin\_rhessys\_input
-   IOin\_tec\_std
-   IOin\_std\_pars
-   IOin\_hdr
-   IOin\_clim
-   IOin\_def\_pars\_simple

### IOin\_rhessys\_input

```{r eval=FALSE}

input_rhessys = IOin_rhessys_input(
  version = file.path("bin", comp_ID, "rhessys7.2"),
  tec_file = "tecfiles/BC.tec",
  world_file = "worldfiles/no_multi/BC_conifer_S_2p.world",
  world_hdr_prefix = "BC",
  flowtable = "flowtables/BC_2p.flow",
  start = "1943 10 01 01",
  end = "1948 10 01 01",
  output_folder = "output/",
  output_prefix = "BC_2p",
  commandline_options = c("-vmort_off -g -b -p -c -msr -v -6")
)

```

### IOin\_tec\_std

```{r eval=FALSE}

tec_std = IOin_tec_std(start = "1943 10 01 01",
                       end = "1948 10 01 01",
                       output_state = TRUE
)

```

### IOin\_std\_pars

```{r eval=FALSE}

std_pars = IOin_std_pars(
  m = 2,
  k = 2,
  m_v = 2,
  k_v = 2,
  pa = 1.15,
  po = 0.766,
  gw1 = 0.24,
  gw2 = 0.2 , 
)

```

### IOin\_hdr

```{r eval=FALSE}

input_hdr_list = IOin_hdr(basin = "defs/basin_p301.def",
                          hillslope = "defs/hill_p301.def",
                          zone = "defs/zone_p301.def",
                          soil = "defs/soil_forestshrub.def",
                          landuse = "defs/lu_p301.def",
                          stratum = c("defs/veg_p301_shrub.def", "defs/veg_p301_conifer.def"),
                          basestations = "clim/upGGmod.base")

```

### Run RHESSys

Together, these inputs (IOin\_X) can be used to run RHESSys for a single simulation with very simple options.

```{r eval=FALSE}

# this is a nearly barebones run
run_rhessys_core(
  input_rhessys = input_rhessys,
  hdr_files = input_hdr_list,
  std_pars = input_std_pars,
  tec_data = input_tec_data
)

```

### IOin\_clim

```{r eval=FALSE}

# this is a very simple case, just creating a basestation
input_clim = IOin_clim(
  base_station_id = 101,
  x_coordinate = 100.0,
  y_coordinate = 100.0,
  z_coordinate = 1984,
  effective_lai = 3.5,
  screen_height = 160,
  daily_prefix = "clim/upGGmod"
)

```

### IOin\_def\_pars

```{r eval=FALSE}

input_def_pars = IOin_def_pars_simple(
  list("defs/veg_p301_conifer.def", "epc.allocation_flag","dickenson"),
  list("defs/veg_p301_conifer.def", "epc.alloc_frootc_leafc", 1),
  list("defs/veg_p301_conifer.def", "epc.alloc_stemc_leafc", 0.6),
  list("defs/veg_p301_conifer.def", "epc.netpabs_shade", 0.2),
  list("defs/soil_forestshrub.def", "soil_depth", 1),
  list("defs/soil_forestshrub.def", "pore_size_index", 0.15),
  list("defs/soil_forestshrub.def", "sat_to_gw_coeff", 0.12),
  list("defs/soil_forestshrub.def", "psi_air_entry", 0.15)
)

```

### Output Variables

```{r eval=FALSE}

  # to use native R-based subsetting
  output_method = "r"

  my_vars = data.frame(c("pd", "Qout"),
                       c("pd", "psn"),
                       c("pdg", "lai"),
                       c("pdg", "plantc"),
                       c("cd", "height"))
  
  output_vars = IOin_output_vars(my_vars)

```

### Run RHESSys with more options

This runs RHESSys for a single simulation with a few more options.

```{r eval=FALSE}

output_data = run_rhessys_core(
  input_rhessys = input_rhessys,
  hdr_files = input_hdr_list,
  def_pars = input_def_pars,
  std_pars = input_std_pars,
  tec_data = input_tec_data,
  clim_base = input_clim,
  output_method = output_method,
  output_variables = output_vars,
  return_data = T
)

```
