---
title: "RHESSysIOinR Utilities Examples"
author: "Will Burke"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## RHESSysIOinR Utilities

The following are various utility functions included with RHESSysIOinR

* `build_redefine()`: generates redefine worldfiles
* `cal.wyd()`: calculates wateryear day
* `change_def_file()`: create new def file based on existing file and input pars
* `clim_auto()`: helper to simplify climate input for `run_rhessys()`
* `evaluation()`: used to evaluate model output
* `input_tec()`: helper to simplify input of tec events
* `mkdate()`: generates wy, yd, and wyd based on day, month, year
* `read_clim()`: reads rhessys formatted climate data into R
* `readin_rhessys_output()`: reads rhessys output into R, includes some subsetting and simply aggregation
* `readin_rhessys_output_old()`: legacy method to read in rhessys output
* `tec_repeat()`: generates repeating tec events for input into `run_rhessys()`, can also copy+rename a redefine file as needed
* `watbal_basin()`: Calculate water balance for a the basin
* `watbal_patch()`: Calculate water balance for a single patch
* `watbal_patch_multi()`: Calculate water balance for a multiple patches
* `write_sample_clim()`: generate new climate based on existing climate

## update_world

```{r }

```
